---
title: "Model"
author: "Cassandra F. Doll, Sarah J. Converse, Collin B. Edwards, and Cheryl B. Schultz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

# Setup
## Loading Required R packages

```{r, echo = T, results = 'hide', message = F, warning = F}
library(lme4)
library(tidyr)
library(emmeans)
library(MASS)
library(here) # Organization
library(ggplot2) # Plotting
library(knitr) # Making table
library(kableExtra) # Making table
set.seed(1) # Arbitrary seed
```

## Function

```{r}
# This function takes our best estimates and standard errors for each parameter in our model (and the estimated environmental variance), and the number of instances to simulate ("Sims") and the number of years to estimate per simulation
# from these it generates `Sims` number of parameter combinations 
# and calculates the expected growth rate for each of the parameter combinations (incorporates uncertainty in parameter estimates)
# It then generates `Years` number of years of realized growth rates for each of these expected growth rates (incorporates environmental variation on top of uncertainty in parameter estimates)
library(boot)
OSBPop <- function(Years,  # Number of years to simulate (we use 1 in all cases)
                   Sims,   # Number of simulations
                   s1.est, # Estimate of survival from egg to pre-diapause larvae, on logit scale
                   s1.se,  # Standard error of survival from egg to pre-diapause larvae, on logit scale
                   s2.est, # Estimate of survival from pre- to post-diapause larvae, on logit scale
                   s2.se,  # Standard error of survival from pre- to post-diapause larvae, on logit scale
                   s3.est, # Estimate of survival from larvae to pupae, on logit scale
                   s3.se,  # Standard error of survival from larvae to pupae, on logit scale
                   s4.est, # Estimate of survival from pupae to adults, on logit scale
                   s4.se,  # Standard error of survival from pupae to adults, on logit scale
                   f.est,  # Estimate of fecundity
                   f.se,   # Standard error of fecundity
                   b.est,  # Estimate of indirect effect
                   b.se,   # Standard error of indirect effect
                   e.sd)   # Standard deviation of year-to-year variability
{ 
  # Generate a matrix to store realized growth rates
  PopLam = matrix(-99,Sims,Years) # Precautionary error check - any -99s after simulation represent some problem with the code
  
  # Generate expected parameters using our estimates for parameter distributions
  # Reminder: survival should be in the range [0, 1], so we use estimates on logit scale
  #   to generate random normal variables in logit space, then transform back to probability
  Parms = data.frame(s1 = inv.logit(rnorm(Sims,s1.est,s1.se)),
                     s2 = inv.logit(rnorm(Sims,s2.est,s2.se)),
                     s3 = inv.logit(rnorm(Sims,s3.est,s3.se)),
                     s4 = inv.logit(rnorm(Sims,s4.est,s4.se)),
                     f = rnorm(Sims,f.est,f.se),
                     b = rnorm(Sims,b.est,b.se))
  # To make sure f is never negative:
  Parms$f[Parms$f<.Machine$double.eps]=.Machine$double.eps
  
  # Combine parameters via our model to calculated expected lambdas 
  exp.lam = Parms$s1 * 
    Parms$s2 *
    Parms$s3 *
    Parms$s4 *
    Parms$f  *
    Parms$b *
    1/2 # Accounting for the fact that half of all individuals are females
  
  # Loop over each simulation and generate realized lambdas using our expected lambdas
  #   (`exp.lam`) and the estimated year to year variation
  # Note that our lambdas are the exponentiation of a log-normal distribution
  #   and this means that the expectation scales with variance
  # We reverse this scaling with the additional term -(e.se^2)/2
  # This can be calculated by hand from the known mean of the lognormal
  for(t in 1:Sims){
    # Take out each r and apply them to each draw
    lam.vec = rnorm(Years,
                    mean = log(exp.lam[t])-(e.sd^2)/2,
                    sd = e.sd)
    # Convert r to lambda
    PopLam[t,] = exp(lam.vec)
  } # Close Sims 
  return(PopLam)
} # Close function
```

# Parameter Estimates
## Years and Sims

```{r}
Years <- 1
Sims <- 10000
```

## s1.est and s1.se

Survival from egg to pre-diapause larvae, data from Doll et al. (**in press**)

```{r}
# Input data
s1.d <- read.csv(here("Data","2018 Herbicide Exposure - Offspring Survival & Development Time.csv"))
s1.d$Treat_ID <- factor(s1.d$Treat_ID,levels = c("U","A","N","C","CA","CN","F","FA","FN"))
# Set U as the reference for Treat_ID
s1.d$Treat_ID <- relevel(as.factor(s1.d$Treat_ID), ref = "U")
# Reduce the dataset
s1.d <- s1.d[,c(2,3,4)]
# Fit a GLMM
s1.m <- glmer(EL_Survi ~ Treat_ID + (1|Fem_ID), family = binomial, data = s1.d)
# Call means and standard errors
s1.p <- as.data.frame(emmeans(s1.m, "Treat_ID"))
## U
s1.est.U <- s1.p$emmean[s1.p$Treat_ID=="U"]
s1.se.U <- s1.p$SE[s1.p$Treat_ID=="U"]
## CA
s1.est.CA <- s1.p$emmean[s1.p$Treat_ID=="CA"]
s1.se.CA <- s1.p$SE[s1.p$Treat_ID=="CA"]
## CN
s1.est.CN <- s1.p$emmean[s1.p$Treat_ID=="CN"]
s1.se.CN <- s1.p$SE[s1.p$Treat_ID=="CN"]
## FA
s1.est.FA <- s1.p$emmean[s1.p$Treat_ID=="FA"]
s1.se.FA <- s1.p$SE[s1.p$Treat_ID=="FA"]
## FN
s1.est.FN <- s1.p$emmean[s1.p$Treat_ID=="FN"]
s1.se.FN <- s1.p$SE[s1.p$Treat_ID=="FN"]
```

## s2.est and s2.se

Survival from pre- to post-diapause larvae, data from Doll et al. (**in press**)

```{r}
# Input data
s2.d <- read.csv(here("Data","2019 Oviposition Behavior - Offspring Survival.csv"))
# Fit a GLMM
s2.m <- glmer(D_Survi ~ (1|Fem_ID), family = binomial, data = s2.d)
# Call means and standard errors
s2.p <- as.data.frame(emmeans(s2.m, "1"))
s2.est <- s2.p[1,2]
s2.se <- s2.p[1,3]
```

## s3.est and s3.se

Survival from larvae to pupae, data from Doll et al. (**in press*)

```{r}
# Input data
s3.d <- read.csv(here("Data","2018 Herbicide Exposure.csv"))
s3.d$Treat_ID <- factor(s3.d$Treat_ID,levels = c("U","A","N","C","CA","CN","F","FA","FN"))
# Set U as the reference for Treat_ID
s3.d$Treat_ID <- relevel(as.factor(s3.d$Treat_ID),ref ="U")
# Remove U03 and U13
U03 <- which(s3.d$Indiv_ID=="U03")
U13 <- which(s3.d$Indiv_ID=="U13")
s3.d <- s3.d[-c(U03,U13),]
# Reduce the dataset
s3.d <- s3.d[,c(1,2,3,4,6)]
# Fit a GLM
s3.m <- glm(LP_Survi ~ Treat_ID + Treat_Date*L_Mass, family = binomial, data = s3.d)
# Call means and standard errors
s3.p <- as.data.frame(emmeans(s3.m, "Treat_ID"))
## U
s3.est.U <- s3.p$emmean[s3.p$Treat_ID=="U"]
s3.se.U <- s3.p$SE[s3.p$Treat_ID=="U"]
## CA
s3.est.CA <- s3.p$emmean[s3.p$Treat_ID=="CA"]
s3.se.CA <- s3.p$SE[s3.p$Treat_ID=="CA"]
## CN
s3.est.CN <- s3.p$emmean[s3.p$Treat_ID=="CN"]
s3.se.CN <- s3.p$SE[s3.p$Treat_ID=="CN"]
## FA
s3.est.FA <- s3.p$emmean[s3.p$Treat_ID=="FA"]
s3.se.FA <- s3.p$SE[s3.p$Treat_ID=="FA"]
## FN
s3.est.FN <- s3.p$emmean[s3.p$Treat_ID=="FN"]
s3.se.FN <- s3.p$SE[s3.p$Treat_ID=="FN"]
```

## s4.est and s4.se

Survival from pupae to adults, data from Doll et al. (**in press**)

```{r}
# Input data
s4.d <- read.csv(here("Data","2018 Herbicide Exposure.csv"))
s4.d$Treat_ID <- factor(s4.d$Treat_ID,levels = c("U","A","N","C","CA","CN","F","FA","FN"))
# Set U as the reference for Treat_ID
s4.d$Treat_ID <- relevel(as.factor(s4.d$Treat_ID),ref = "U")
# Remove U03 and U13
U03 <- which(s4.d$Indiv_ID=="U03")
U13 <- which(s4.d$Indiv_ID=="U13")
s4.d <- s4.d[-c(U03,U13),]
# Reduce the dataset
s4.d <- s4.d[,c(1,2,3,5,6)]
# Fit a GLM
s4.m <- glm(PE_Survi ~ Treat_ID + Treat_Date*L_Mass, family = binomial, data = s4.d)
# Call means and standard errors
s4.p <- as.data.frame(emmeans(s4.m, "Treat_ID"))
## U
s4.est.U <- s4.p[1,2]
## For U's standard error: average the standard errors of the other treatments
s4.se.U <- s4.p$SE
s4.se.U <- s4.se.U[-1]
s4.se.U <- mean(s4.se.U)
## CA
s4.est.CA <- s4.p$emmean[s4.p$Treat_ID=="CA"]
s4.se.CA <- s4.p$SE[s4.p$Treat_ID=="CA"]
## CN
s4.est.CN <- s4.p$emmean[s4.p$Treat_ID=="CN"]
s4.se.CN <- s4.p$SE[s4.p$Treat_ID=="CN"]
## FA
s4.est.FA <- s4.p$emmean[s4.p$Treat_ID=="FA"]
s4.se.FA <- s4.p$SE[s4.p$Treat_ID=="FA"]
## FN
s4.est.FN <- s4.p$emmean[s4.p$Treat_ID=="FN"]
s4.se.FN <- s4.p$SE[s4.p$Treat_ID=="FN"]
```

## f.est and f.se

Fecundity, data from Doll et al. (**in press**)

```{r}
# Input data
f.d <- read.csv(here("Data","2018 Herbicide Exposure.csv"))
f.d$Treat_ID <- factor(f.d$Treat_ID,levels = c("U","A","N","C","CA","CN","F","FA","FN"))
# Set U as the reference for Treat_ID
f.d$Treat_ID <- relevel(as.factor(f.d$Treat_ID),ref = "U")
# Remove females that died before laying any eggs
C07 <- which(f.d$Indiv_ID=="C07")
C40 <- which(f.d$Indiv_ID=="C40")
CN24 <- which(f.d$Indiv_ID=="CN24")
N21 <- which(f.d$Indiv_ID=="N21")
N40 <- which(f.d$Indiv_ID=="N40")
f.d <- f.d[-c(C07,C40,CN24,N21,N40),]
# Reduce the dataset
f.d <- f.d[,c(1,2,16)]
f.d <- na.omit(f.d)
# Fit a negative binomial GLM
f.m <- glm.nb(Fem_Fecun ~ Treat_ID, link = log, data = f.d)
d <- f.m$deviance/f.m$df.residual
# Call means and standard errors
f.p <- as.data.frame(emmeans(f.m, type = "response", "Treat_ID"))
## U
f.est.U <- f.p$response[f.p$Treat_ID=="U"]
f.se.U <- f.p$SE[f.p$Treat_ID=="U"]
## CA
f.est.CA <- f.p$response[f.p$Treat_ID=="CA"]
f.se.CA <- f.p$SE[f.p$Treat_ID=="CA"]
## CN
f.est.CN <- f.p$response[f.p$Treat_ID=="CN"]
f.se.CN <- f.p$SE[f.p$Treat_ID=="CN"]
## FA
f.est.FA <- f.p$response[f.p$Treat_ID=="FA"]
f.se.FA <- f.p$SE[f.p$Treat_ID=="FA"]
## FN
f.est.FN <- f.p$response[f.p$Treat_ID=="FN"]
f.se.FN <- f.p$SE[f.p$Treat_ID=="FN"]
```

## b.est and b.se

Estimates of indirect effects (benefit of herbicides to caterpillar due to reduced competition of plants), from A. B. James, Washington State University (WSU) 

```{r}
# Input data
b.d <- read.csv(here("Data","2019 A. B. James.csv"))
# Subset VC and PC
b.d.NC <- b.d[b.d$Treat_ID=="NC",]
b.d.C <- b.d[b.d$Treat_ID=="C",]
# Vectors for A_Mass
NC <- b.d.NC$A_Mass
C <- b.d.C$A_Mass
# Bootstrap
boots <- 100000
m <- rep(NA,boots)
b.est <- (mean(NC)-mean(C))/mean(C) + 1

for(i in 1:boots){
  NC.boot <- sample(NC,length(NC),replace=T)
  C.boot <- sample(C,length(C),replace=T)
  m[i] <- abs((mean(NC.boot)-mean(C.boot))/mean(C.boot))+1
}
(b.se <- sd(m))
(ci <- quantile(m,probs = c(0.025,0.975)))
```

## e.sd

Estimating environmental variation in growth rate, data from USFWS annual index counts for Oregon silverspot abundance at Mt Hebo

```{r}
# Input data
e.d <- read.csv(here("Data","1990-2002 Index Counts.csv"))
# Calculate the annual growth rate at Mount_Hebo
r <- log(e.d$Mt_Hebo[-1]/e.d$Mt_Hebo[-length(e.d$Mt_Hebo)])
# Calculate the standard deviation of the annual growth rate
e.sd <- sd(r)
```

# Results {.tabset}

```{r}
# Create a data frame to store a summary of the results
res.df <- data.frame(treat = c("U","CA.D","CA.DI","CN.D","CN.DI","FA.D","FA.DI","FN.D","FN.DI"),
                     mean = rep(-99,9),
                     sd = rep(-99,9),
                     ci.025 = rep(-99,9),
                     ci.16 = rep(-99,9),
                     ci.84 = rep(-99,9),
                     ci.975 = rep(-99,9),
                     RGR = rep(-99,9))
# This will become a matrix/data frame of all bootstrapped means
mean.df <- NULL
```

## U

Reminder: Since there is no indirect benefit, direct and indirect effects are the same

```{r}
U <- OSBPop(Years,             # Number of years
            Sims,              # Number of simulations
            s1.est = s1.est.U, # Estimate of survival from egg to pre-diapause larvae
            s1.se = s1.se.U,   # Standard error of survival from egg to pre-diapause larvae
            s2.est = s2.est,   # Estimate of survival from pre- to post-diapause larvae
            s2.se = s2.se,     # Standard error of survival from pre- to post-diapause larvae
            s3.est = s3.est.U, # Estimate of survival from larvae to pupae
            s3.se = s3.se.U,   # Standard error of survival from larvae to pupae
            s4.est = s4.est.U, # Estimate of survival from pupae to adults
            s4.se = s4.se.U,   # Standard error of survival from pupae to adults
            f.est = f.est.U,   # Estimate of fecundity
            f.se = f.se.U,     # Standard error of fecundity
            b.est = 1,         # Estimate of indirect effect
            b.se = 0,          # Standard error of indirect effect
            e.sd)              # Standard deviation of year-to-year variability

mean.U <- rowMeans(U)
# Save
mean.df <- cbind(mean.df,mean.U)

print("Mean:")
mean(mean.U)
# Save
res.df$mean[res.df$treat=="U"] <- mean(mean.U)

print("SE/SD:")
sd(mean.U)
# Save
res.df$sd[res.df$treat=="U"] <- sd(mean.U)

print("95% CI:")
quantile(mean.U,c(0.025,0.975))
# Save
res.df[res.df$treat=="U",c("ci.025","ci.975")] <-
  quantile(mean.U,c(0.025,0.975))

print("+/- 1SE:")
quantile(mean.U,c(pnorm(-1),pnorm(1)))
# Save
res.df[res.df$treat=="U",c("ci.16","ci.84")] <-
  quantile(mean.U,c(pnorm(-1),pnorm(1)))

print("RGR: 1 (by definition), CI = [1,1] (by definition)")
res.df$RGR[res.df$treat=="U"] <- 1

hist(mean.U,breaks = 50)
```

## CA

### Direct Effects Only

```{r}
# Assuming only direct effects
CA.D <- OSBPop(Years,              # Number of years
               Sims,               # Number of simulations
               s1.est = s1.est.CA, # Estimate of survival from egg to pre-diapause larvae
               s1.se = s1.se.CA,   # Standard error of survival from egg to pre-diapause larvae
               s2.est = s2.est,    # Estimate of survival from pre- to post-diapause larvae
               s2.se = s2.se,      # Standard error of survival from pre- to post-diapause larvae
               s3.est = s3.est.CA, # Estimate of survival from larvae to pupae
               s3.se = s3.se.CA,   # Standard error of survival from larvae to pupae
               s4.est = s4.est.CA, # Estimate of survival from pupae to adults
               s4.se = s4.se.CA,   # Standard error of survival from pupae to adults
               f.est = f.est.CA,   # Estimate of fecundity
               f.se = f.se.CA,     # Standard error of fecundity
               b.est = 1,          # Estimate of indirect effect
               b.se = 0,           # Standard error of indirect effect
               e.sd)               # Standard deviation of year-to-year variability

mean.CA.D <- rowMeans(CA.D)
# Save
mean.df <- cbind(mean.df,mean.CA.D)

print("Mean:")
mean(mean.CA.D)
# Save
res.df$mean[res.df$treat=="CA.D"] <- mean(mean.CA.D)

print("SE/SD:")
sd(mean.CA.D)
# Save
res.df$sd[res.df$treat=="CA.D"] <- sd(mean.CA.D)

print("95% CI:")
quantile(mean.CA.D,c(0.025,0.975))
# Save
res.df[res.df$treat=="CA.D",c("ci.025","ci.975")] <-
  quantile(mean.CA.D,c(0.025,0.975))

print("+/- 1SE:")
quantile(mean.CA.D,c(pnorm(-1),pnorm(1)))
# Save
res.df[res.df$treat=="CA.D",c("ci.16","ci.84")] <-
  quantile(mean.CA.D,c(pnorm(-1),pnorm(1)))

print("RGR: ")
mean(mean.CA.D)/mean(mean.U)
#Save
res.df$RGR[res.df$treat=="CA.D"] <- mean(mean.CA.D)/mean(mean.U)
print("95% CI: ")
print(quantile(mean.CA.D/mean.U, c(.025, .975)))


hist(mean.CA.D,breaks = 50)
```

### Direct and Indirect Effects

```{r}
# Assuming both direct and indirect effects
CA.DI <- OSBPop(Years,              # Number of years
                Sims,               # Number of simulations
                s1.est = s1.est.CA, # Estimate of survival from egg to pre-diapause larvae
                s1.se = s1.se.CA,   # Standard error of survival from egg to pre-diapause larvae
                s2.est = s2.est,    # Estimate of survival from pre- to post-diapause larvae
                s2.se = s2.se,      # Standard error of survival from pre- to post-diapause larvae
                s3.est = s3.est.CA, # Estimate of survival from larvae to pupae
                s3.se = s3.se.CA,   # Standard error of survival from larvae to pupae
                s4.est = s4.est.CA, # Estimate of survival from pupae to adults
                s4.se = s4.se.CA,   # Standard error of survival from pupae to adults
                f.est = f.est.CA,   # Estimate of fecundity
                f.se = f.se.CA,     # Standard error of fecundity
                b.est,              # Estimate of indirect effect
                b.se,               # Standard error of indirect effect
                e.sd)               # Standard deviation of year-to-year variability

mean.CA.DI <- rowMeans(CA.DI)
# Save
mean.df <- cbind(mean.df,mean.CA.DI)

print("Mean:")
mean(mean.CA.DI)
# Save
res.df$mean[res.df$treat=="CA.DI"] <- mean(mean.CA.DI)

print("SE/SD:")
sd(mean.CA.DI)
# Save
res.df$sd[res.df$treat=="CA.DI"] <- sd(mean.CA.DI)

print("95% CI:")
quantile(mean.CA.DI,c(0.025,0.975))
# Save
res.df[res.df$treat=="CA.DI",c("ci.025","ci.975")] <-
  quantile(mean.CA.DI,c(0.025,0.975))

print("+/- 1SE:")
quantile(mean.CA.DI,c(pnorm(-1),pnorm(1)))
# Save
res.df[res.df$treat=="CA.DI",c("ci.16","ci.84")] <-
  quantile(mean.CA.DI,c(pnorm(-1),pnorm(1)))

print("RGR: ")
mean(mean.CA.DI)/mean(mean.U)
#Save
res.df$RGR[res.df$treat=="CA.DI"] <- mean(mean.CA.DI)/mean(mean.U)
print("95% CI: ")
print(quantile(mean.CA.DI/mean.U, c(.025, .975)))



hist(mean.CA.DI,breaks = 50)
```

## CN

### Direct Effects Only

```{r}
# Assuming only direct effects
CN.D <- OSBPop(Years,              # Number of years
               Sims,               # Number of simulations
               s1.est = s1.est.CN, # Estimate of survival from egg to pre-diapause larvae
               s1.se = s1.se.CN,   # Standard error of survival from egg to pre-diapause larvae
               s2.est = s2.est,    # Estimate of survival from pre- to post-diapause larvae
               s2.se = s2.se,      # Standard error of survival from pre- to post-diapause larvae
               s3.est = s3.est.CN, # Estimate of survival from larvae to pupae
               s3.se = s3.se.CN,   # Standard error of survival from larvae to pupae
               s4.est = s4.est.CN, # Estimate of survival from pupae to adults
               s4.se = s4.se.CN,   # Standard error of survival from pupae to adults
               f.est = f.est.CN,   # Estimate of fecundity
               f.se = f.se.CN,     # Standard error of fecundity
               b.est = 1,          # Estimate of indirect effect
               b.se = 0,           # Standard error of indirect effect
               e.sd)               # Standard deviation of year-to-year variability

mean.CN.D <- rowMeans(CN.D)
# Save
mean.df <- cbind(mean.df,mean.CN.D)

print("Mean:")
mean(mean.CN.D)
# Save
res.df$mean[res.df$treat=="CN.D"] <- mean(mean.CN.D)

print("SE/SD:")
sd(mean.CN.D)
# Save
res.df$sd[res.df$treat=="CN.D"] <- sd(mean.CN.D)

print("95% CI:")
quantile(mean.CN.D,c(0.025,0.975))
# Save
res.df[res.df$treat=="CN.D",c("ci.025","ci.975")] <-
  quantile(mean.CN.D,c(0.025,0.975))

print("+/- 1SE:")
quantile(mean.CN.D,c(pnorm(-1),pnorm(1)))
# Save
res.df[res.df$treat=="CN.D",c("ci.16","ci.84")] <-
  quantile(mean.CN.D,c(pnorm(-1),pnorm(1)))

print("RGR: ")
mean(mean.CN.D)/mean(mean.U)
#save
res.df$RGR[res.df$treat=="CN.D"] <- mean(mean.CN.D)/mean(mean.U)
print("95% CI: ")
print(quantile(mean.CN.D/mean.U, c(.025, .975)))



hist(mean.CN.D,breaks = 50)
```

### Direct and Indirect Effects

```{r}
# Assuming both direct and indirect effects
CN.DI <- OSBPop(Years,              # Number of years
                Sims,               # Number of simulations
                s1.est = s1.est.CN, # Estimate of survival from egg to pre-diapause larvae
                s1.se = s1.se.CN,   # Standard error of survival from egg to pre-diapause larvae
                s2.est = s2.est,    # Estimate of survival from pre- to post-diapause larvae
                s2.se = s2.se,      # Standard error of survival from pre- to post-diapause larvae
                s3.est = s3.est.CN, # Estimate of survival from larvae to pupae
                s3.se = s3.se.CN,   # Standard error of survival from larvae to pupae
                s4.est = s4.est.CN, # Estimate of survival from pupae to adults
                s4.se = s4.se.CN,   # Standard error of survival from pupae to adults
                f.est = f.est.CN,   # Estimate of fecundity
                f.se = f.se.CN,     # Standard error of fecundity
                b.est,              # Estimate of indirect effect
                b.se,               # Standard error of indirect effect
                e.sd)               # Standard deviation of year-to-year variability

mean.CN.DI <- rowMeans(CN.DI)
# Save
mean.df <- cbind(mean.df,mean.CN.DI)

print("Mean:")
mean(mean.CN.DI)
# Save
res.df$mean[res.df$treat=="CN.DI"] <- mean(mean.CN.DI)

print("SE/SD:")
sd(mean.CN.DI)
# Save
res.df$sd[res.df$treat=="CN.DI"] <- sd(mean.CN.DI)

print("95% CI:")
quantile(mean.CN.DI,c(0.025,0.975))
# Save
res.df[res.df$treat=="CN.DI",c("ci.025","ci.975")] <-
  quantile(mean.CN.DI,c(0.025,0.975))

print("+/- 1SE:")
quantile(mean.CN.DI,c(pnorm(-1),pnorm(1)))
# Save
res.df[res.df$treat=="CN.DI",c("ci.16","ci.84")] <-
  quantile(mean.CN.DI,c(pnorm(-1),pnorm(1)))

print("RGR: ")
mean(mean.CN.DI)/mean(mean.U)
#save
res.df$RGR[res.df$treat=="CN.DI"] <- mean(mean.CN.DI)/mean(mean.U)
print("95% CI: ")
print(quantile(mean.CN.DI/mean.U, c(.025, .975)))



hist(mean.CN.DI,breaks = 50)
```

## FA

### Direct Effects Only

```{r}
# Assuming only direct effects
FA.D <- OSBPop(Years,              # Number of years
               Sims,               # Number of simulations
               s1.est = s1.est.FA, # Estimate of survival from egg to pre-diapause larvae
               s1.se = s1.se.FA,   # Standard error of survival from egg to pre-diapause larvae
               s2.est = s2.est,    # Estimate of survival from pre- to post-diapause larvae
               s2.se = s2.se,      # Standard error of survival from pre- to post-diapause larvae
               s3.est = s3.est.FA, # Estimate of survival from larvae to pupae
               s3.se = s3.se.FA,   # Standard error of survival from larvae to pupae
               s4.est = s4.est.FA, # Estimate of survival from pupae to adults
               s4.se = s4.se.FA,   # Standard error of survival from pupae to adults
               f.est = f.est.FA,   # Estimate of fecundity
               f.se = f.se.FA,     # Standard error of fecundity
               b.est = 1,          # Estimate of indirect effect
               b.se = 0,           # Standard error of indirect effect
               e.sd)               # Standard deviation of year-to-year variability

mean.FA.D <- rowMeans(FA.D)
# Save
mean.df <- cbind(mean.df,mean.FA.D)

print("Mean:")
mean(mean.FA.D)
# Save
res.df$mean[res.df$treat=="FA.D"] <- mean(mean.FA.D)

print("SE/SD:")
sd(mean.FA.D)
# Save
res.df$sd[res.df$treat=="FA.D"] <- sd(mean.FA.D)

print("95% CI:")
quantile(mean.FA.D,c(0.025,0.975))
# Save
res.df[res.df$treat=="FA.D",c("ci.025","ci.975")] <-
  quantile(mean.FA.D,c(0.025,0.975))

print("+/- 1SE:")
quantile(mean.FA.D,c(pnorm(-1),pnorm(1)))
# Save
res.df[res.df$treat=="FA.D",c("ci.16","ci.84")] <-
  quantile(mean.FA.D,c(pnorm(-1),pnorm(1)))

print("RGR: ")
mean(mean.FA.D)/mean(mean.U)
#save
res.df$RGR[res.df$treat=="FA.D"] <- mean(mean.FA.D)/mean(mean.U)
print("95% CI: ")
print(quantile(mean.FA.D/mean.U, c(.025, .975)))



hist(mean.FA.D,breaks = 50)
```

### Direct and Indirect Effects

```{r}
# Assuming both direct and indirect effects
FA.DI <- OSBPop(Years,              # Number of years
                Sims,               # Number of simulations
                s1.est = s1.est.FA, # Estimate of survival from egg to pre-diapause larvae
                s1.se = s1.se.FA,   # Standard error of survival from egg to pre-diapause larvae
                s2.est = s2.est,    # Estimate of survival from pre- to post-diapause larvae
                s2.se = s2.se,      # Standard error of survival from pre- to post-diapause larvae
                s3.est = s3.est.FA, # Estimate of survival from larvae to pupae
                s3.se = s3.se.FA,   # Standard error of survival from larvae to pupae
                s4.est = s4.est.FA, # Estimate of survival from pupae to adults
                s4.se = s4.se.FA,   # Standard error of survival from pupae to adults
                f.est = f.est.FA,   # Estimate of fecundity
                f.se = f.se.FA,     # Standard error of fecundity
                b.est,              # Estimate of indirect effect
                b.se,               # Standard error of indirect effect
                e.sd)               # Standard deviation of year-to-year variability

mean.FA.DI <- rowMeans(FA.DI)
# Save
mean.df <- cbind(mean.df,mean.FA.DI)

print("Mean:")
mean(mean.FA.DI)
# Save
res.df$mean[res.df$treat=="FA.DI"] <- mean(mean.FA.DI)

print("SE/SD:")
sd(mean.FA.DI)
# Save
res.df$sd[res.df$treat=="FA.DI"] <- sd(mean.FA.DI)

print("95% CI:")
quantile(mean.FA.DI,c(0.025,0.975))
# Save
res.df[res.df$treat=="FA.DI",c("ci.025","ci.975")] <-
  quantile(mean.FA.DI,c(0.025,0.975))

print("+/- 1SE:")
quantile(mean.FA.DI,c(pnorm(-1),pnorm(1)))
# Save
res.df[res.df$treat=="FA.DI",c("ci.16","ci.84")] <-
  quantile(mean.FA.DI,c(pnorm(-1),pnorm(1)))

print("RGR: ")
mean(mean.FA.DI)/mean(mean.U)
#save
res.df$RGR[res.df$treat=="FA.DI"] <- mean(mean.FA.DI)/mean(mean.U)
print("95% CI: ")
print(quantile(mean.FA.DI/mean.U, c(.025, .975)))


hist(mean.FA.DI,breaks = 50)
```

## FN

### Direct Effects Only

```{r}
# Assuming only direct effects
FN.D <- OSBPop(Years,              # Number of years
               Sims,               # Number of simulations
               s1.est = s1.est.FN, # Estimate of survival from egg to pre-diapause larvae
               s1.se = s1.se.FN,   # Standard error of survival from egg to pre-diapause larvae
               s2.est = s2.est,    # Estimate of survival from pre- to post-diapause larvae
               s2.se = s2.se,      # Standard error of survival from pre- to post-diapause larvae
               s3.est = s3.est.FN, # Estimate of survival from larvae to pupae
               s3.se = s3.se.FN,   # Standard error of survival from larvae to pupae
               s4.est = s4.est.FN, # Estimate of survival from pupae to adults
               s4.se = s4.se.FN,   # Standard error of survival from pupae to adults
               f.est = f.est.FN,   # Estimate of fecundity
               f.se = f.se.FN,     # Standard error of fecundity
               b.est = 1,          # Estimate of indirect effect
               b.se = 0,           # Standard error of indirect effect
               e.sd)               # Standard deviation of year-to-year variability

mean.FN.D <- rowMeans(FN.D)
# Save
mean.df <- cbind(mean.df,mean.FN.D)

print("Mean:")
mean(mean.FN.D)
# Save
res.df$mean[res.df$treat=="FN.D"] <- mean(mean.FN.D)

print("SE/SD:")
sd(mean.FN.D)
# Save
res.df$sd[res.df$treat=="FN.D"] <- sd(mean.FN.D)

print("95% CI:")
quantile(mean.FN.D,c(0.025,0.975))
# Save
res.df[res.df$treat=="FN.D",c("ci.025","ci.975")] <-
  quantile(mean.FN.D,c(0.025,0.975))

print("+/- 1SE:")
quantile(mean.FN.D,c(pnorm(-1),pnorm(1)))
# Save
res.df[res.df$treat=="FN.D",c("ci.16","ci.84")] <-
  quantile(mean.FN.D,c(pnorm(-1),pnorm(1)))

print("RGR: ")
mean(mean.FN.D)/mean(mean.U)
#save
res.df$RGR[res.df$treat=="FN.D"] <- mean(mean.FN.D)/mean(mean.U)
print("95% CI: ")
print(quantile(mean.FN.D/mean.U, c(.025, .975)))



hist(mean.FN.D,breaks = 50)
```

### Direct and Indirect Effects

```{r}
# Assuming both direct and indirect effects
FN.DI <- OSBPop(Years,              # Number of years
                Sims,               # Number of simulations
                s1.est = s1.est.FN, # Estimate of survival from egg to pre-diapause larvae
                s1.se = s1.se.FN,   # Standard error of survival from egg to pre-diapause larvae
                s2.est = s2.est,    # Estimate of survival from pre- to post-diapause larvae
                s2.se = s2.se,      # Standard error of survival from pre- to post-diapause larvae
                s3.est = s3.est.FN, # Estimate of survival from larvae to pupae
                s3.se = s3.se.FN,   # Standard error of survival from larvae to pupae
                s4.est = s4.est.FN, # Estimate of survival from pupae to adults
                s4.se = s4.se.FN,   # Standard error of survival from pupae to adults
                f.est = f.est.FN,   # Estimate of fecundity
                f.se = f.se.FN,     # Standard error of fecundity
                b.est,              # Estimate of indirect effect
                b.se,               # Standard error of indirect effect
                e.sd)               # Standard deviation of year-to-year variability

mean.FN.DI <- rowMeans(FN.DI)
# Save
mean.df <- cbind(mean.df,mean.FN.DI)

print("Mean:")
mean(mean.FN.DI)
# Save
res.df$mean[res.df$treat=="FN.DI"] <- mean(mean.FN.DI)

print("SE/SD:")
sd(mean.FN.DI)
# Save
res.df$sd[res.df$treat=="FN.DI"] <- sd(mean.FN.DI)

print("95% CI:")
quantile(mean.FN.DI,c(0.025,0.975))
# Save
res.df[res.df$treat=="FN.DI",c("ci.025","ci.975")] <-
  quantile(mean.FN.DI,c(0.025,0.975))

print("+/- 1SE:")
quantile(mean.FN.DI,c(pnorm(-1),pnorm(1)))
# Save
res.df[res.df$treat=="FN.DI",c("ci.16","ci.84")] <-
  quantile(mean.FN.DI,c(pnorm(-1),pnorm(1)))

print("RGR: ")
mean(mean.FN.DI)/mean(mean.U)
#save
res.df$RGR[res.df$treat=="FN.DI"] <- mean(mean.FN.DI)/mean(mean.U)
print("95% CI: ")
print(quantile(mean.FN.DI/mean.U, c(.025, .975)))


hist(mean.FN.DI,breaks = 50)
```

## Save

```{r}
# Change mean.df into a data frame
mean.df <- as.data.frame(mean.df)
# Remove "mean" from labels
colnames(mean.df) <- gsub("mean.","",colnames(mean.df))
saveRDS(mean.df,here("Results","Raw Means.RDS"))

## Change results columns to have separate column for treatment and for direct vs indirect
res.save = res.df
# For untreated, duplicate rows for direct and indirect (since indirect = direct)
res.save = rbind(res.save[1,], res.save)
# Update to have consistent suffix for Direct vs Indirect - easier gsub later
res.save$treat[1:2]= c("U.D", "U.DI")
res.save$effect = gsub('.*[.]', "", res.save$treat)
res.save$treat = gsub('[.].*', "", res.save$treat)
res.save = res.save %>% 
  dplyr::relocate(effect, .after = treat)
  
write.csv(res.save, here("Results","Model Results.csv"))
```

```{r}
kable(res.save) %>%
   kable_styling(bootstrap_options = "striped", full_width = F)
```

# Figure

```{r}
res.df <- read.csv(here("Results","Model Results.csv"))
res.df$treat <- factor(res.df$treat,levels = c("U","CA","CN","FA","FN"))

Fig_2 <-
  ggplot(res.df, aes(x = treat, y = mean, shape = effect, fill = treat)) +
  geom_errorbar(aes(ymin = ci.025, ymax = ci.975),
                width = 0.3, size = 0.5, position = position_dodge(width = 0.5)) +
  geom_point(size = 3, position = position_dodge(width = 0.5))

Fig_2 <- Fig_2 + 
  scale_shape_manual(name = "Effect", breaks = c("D","DI"), labels = c("Direct", "Direct + Indirect"),
                     values = c(21,24)) + 
  scale_fill_manual(name = "Alternative", breaks = c("U","CA","CN","FA","FN"), 
                    labels = c("Untreated (U)","Clopyralid + Agri-Dex (CA)",
                               "Clopyralid + Nu Film (CN)","Fluazifop-p-butyl + Agri-Dex (FA)",
                               "Fluazifop-p-butyl + Nu Film (FN)"),
                    values = c("#999999","#56B4E9","#009E73","#D55E00","#CC79A7")) + 
  theme(axis.text.x = element_text(size = 8.5), axis.text.y = element_text(size = 8.5), 
        title = element_text(size = 8.5), panel.background = element_blank(), 
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"), axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black")) +
  theme(legend.position = "top", legend.text = element_text(size = 8.5),
        legend.title = element_text(size = 8.5),
        legend.key = element_rect(fill = "white")) +
  guides(fill = guide_legend(override.aes = list(shape = 22), nrow = 3, byrow = FALSE),
         shape = guide_legend(nrow = 2)) +
  scale_y_continuous("Population Growth Rate",
                   breaks = 30*(0:4))+
  xlab("Alternative")
print(Fig_2)

ggsave(
  here("Figures", "Fig. 2.jpg"),
  Fig_2,
  width = 174,
  height = 100,
  units = c("mm"),
  dpi = 600
)
```
